What's hot and what's not:\\Windowed developer topic analysis

### 2009-07-01 lda-paper.tex
Some more edits

Topic analysis 
uses tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
documents (commit log comments)~\cite{1374321-short,10.1109/ICPC.2007.13-short,lukins2008,10.1109/MSR.2007.20-short}.
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS).

Previous work on topic analysis has been applied to the entire history of a 
project~\cite{1321709-short,lukins2008,10.1109/ICPC.2007.13-short}
to produce a single set of topics. In this paper we explore the idea
of time-windowed topic analysis; that is, we pick a period of, say,
one month and perform topic analysis within each window.  In addition
to being able to study just which topics were actually worked on and
when, we can also analyze the historical timeline of topics for
patterns and trends.  For example, we might wish to categorize each
topic as being local to a single time window, spanning multiple
consecutive windows, or recurring periodically.  Alternatively, we
might explore evolutionary patterns, such as if topics often recur together, if one particular
topic usually precedes another, if unplanned topics outnumber planned
topics, if particular topics recur near the end of development
iterations, etc. % whereas this paper suggests that most topics are
%local.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
%(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
(LSI)~\cite{1374321-short,10.1109/ICPC.2007.13-short},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774-short,1566153-short}. These tools are used for
document modelling, document clustering and collaborative
filtering. LDA attempts to encode documents as a mixture
model, a combination of topics.  LDA has been used in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20-short}%,NIPS2007637,1321709}.
to extract topics from documents such as methods, bug
reports, source code and files.

@@

Lukins et al.~\cite{lukins2008} use LDA to help bug
localization by querying for documents that are related to a bug's
topic. They used LDA to build a topic model and then searched for
similar documents by querying with sample documents.
%These queries would indicate the proportion of
%topics that were associated with the query and the
%related documents.

@@

 The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
  attempts to show all topics and trends at once across time in a
  compact view that could fit on one page.   The
\emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
the larger trends to the bottom.  Once the larger trends are stacked
we fill in the gaps with the smaller trends in the same window, and
then stack the smaller trends on top.  Although there is a chance that
gaps will be left due to non-optimal stacking, in practice there are
many small trends (90\% of all trends contain 1 topic) that fill in
these gaps quickly.  Different instances of the same trend share the
same color; apart from that, the color of a trend is randomly chosen
while the color of a unique topic is grey. Color similarity is not meaningful.  
%The compact-trend-view is
%convenient because it shows all of the topic information, as shown by
%the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
%\ref{fig:zoomedsmear}. 
The \emph{compact-trend-view} makes repeating
continuous trends easy to pick out, although discontinuous trends are
harder to spot, and provides a general summary of the topics within a project.

@@

We wish to extend this study by further exploring parameter
  choices and their effects in terms of window overlap size and number
 of topics.
%  We want to explore parameters more across multiple projects,
%   what motivates the need for more or less topics.  Parameters
%   include: window overlapping, window overlap size, window size,
%   number of topics, etc.  We want to investigate the effects of
%   overlapping windows on this analysis. Do they make the trends too
%   long? Do they highlight some trends unnecessarily or are they more
%   because they are time agnostic. 
Another avenue of future work is automatic topic labelling. Given a word
distribution we should be able to automatically select a word or term
that describes that distribution. Potential external sources of topic
names include software engineering taxonomies, ontologies and
standards.
\newline\noindent{\small\textbf{Acknowledgements:} Thanks to Tudor Girba and Neil Ernst.}
%for valuable feedback
\shrinkit
\bibliographystyle{latex8}

>>>>

Topic analysis 
uses tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
documents (commit log comments)~\cite{1374321-short,10.1109/ICPC.2007.13-short,lukins2008-short,10.1109/MSR.2007.20-short}.
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS).

Previous work on topic analysis has been applied to the entire history of a 
project~\cite{1321709-short,lukins2008-short,10.1109/ICPC.2007.13-short}
to produce a single set of topics. In this paper we explore the idea
of time-windowed topic analysis; that is, we pick a period of, say,
one month and perform topic analysis within each window.  In addition
to being able to study just which topics were actually worked on and
when, we can also analyze the historical timeline of topics for
patterns and trends.  For example, we might wish to categorize each
topic as being local to a single time window, spanning multiple
consecutive windows, or recurring periodically.  Alternatively, we
might explore evolutionary patterns, such as if topics often recur together, if one particular
topic usually precedes another, if unplanned topics outnumber planned
topics, if particular topics recur near the end of development
iterations, etc. % whereas this paper suggests that most topics are
%local.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
%(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
(LSI)~\cite{1374321-short,10.1109/ICPC.2007.13-short},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774-short,1566153-short}. These tools are used for
document modelling, document clustering and collaborative
filtering. LDA attempts to encode documents as a mixture
model, a combination of topics.  LDA has been used in software engineering
literature~\cite{lukins2008-short,10.1109/MSR.2007.20-short}%,NIPS2007637,1321709}.
to extract topics from documents such as methods, bug
reports, source code and files.

@@

Lukins et al.~\cite{lukins2008-short} use LDA to help bug
localization by querying for documents that are related to a bug's
topic. They used LDA to build a topic model and then searched for
similar documents by querying with sample documents.
%These queries would indicate the proportion of
%topics that were associated with the query and the
%related documents.

@@

 The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
  attempts to show all topics and trends at once across time in a
  compact view that could fit on one page.   The
\emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
the larger trends to the bottom.  Once the larger trends are stacked
we fill in the gaps with the smaller trends in the same window, and
then stack the smaller trends on top.  Although there is a chance that
gaps will be left due to non-optimal stacking, in practice there are
many small trends (90\% of all trends contain 1 topic) that fill in
these gaps quickly.  Different instances of the same trend share the
same color; apart from that, the color of a trend is randomly chosen
while topics that do not reoccur are colored grey.
%the color of a unique topic is grey. 
Color similarity is not meaningful.  
%The compact-trend-view is
%convenient because it shows all of the topic information, as shown by
%the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
%\ref{fig:zoomedsmear}. 
The \emph{compact-trend-view} makes repeating
continuous trends easy to pick out, although discontinuous trends are
harder to spot, and provides a general summary of the topics within a project.

@@

We wish to extend this study by further exploring parameter
  choices and their effects in terms of window overlap size and number
 of topics.
%  We want to explore parameters more across multiple projects,
%   what motivates the need for more or less topics.  Parameters
%   include: window overlapping, window overlap size, window size,
%   number of topics, etc.  We want to investigate the effects of
%   overlapping windows on this analysis. Do they make the trends too
%   long? Do they highlight some trends unnecessarily or are they more
%   because they are time agnostic. 
Another avenue of future work is automatic topic labelling. Given a word
distribution we should be able to automatically select a word or term
that describes that distribution. Potential external sources of topic
names include software engineering taxonomies, ontologies and
standards.
\newline\noindent{\small\textbf{Acknowledgements:} Thanks to Tudor Girba and Neil Ernst.}
%for their valuable feedback.}
\shrinkit
\bibliographystyle{latex8}
